---
title: "Correlations between CT and clinical parameters"
output: html_notebook
---


Get the data.
```{r}
raw_data <- read.csv2("C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/FD_data/Expi.csv")
raw_data2 <- read.csv("C:/Users/csuehs/Dropbox/SCANN'AIR_princeps/data-verification/Data_scannair_NA_corrections_isaetcarey_20170427.csv")
```

Extract the appropriate variables. We are interested in raw FD values for attentuation at -850HU and according to two factors: 

* inspiratory versus expiratory measures;
* upper versus lower lung.
      ---slice 1
      ---slice 10
```{r}

df_fd <- data.frame(raw_data$numpat,
                  raw_data$I.850.H1,
                  raw_data$I.850.H10,
                  raw_data$I.850.L1,
                  raw_data$I.850.L10,
                  raw_data$E.850.H1,
                  raw_data$E.850.H10,
                  raw_data$E.850.L1,
                  raw_data$E.850.L10)

colnames(df_fd) <- c("patient",
                   "fd_i_up_1", 
                    "fd_i_up_10", 
                    "fd_i_lo_1", 
                    "fd_i_lo_10", 
                    "fd_e_up_1", 
                    "fd_e_up_10", 
                    "fd_e_lo_1", 
                    "fd_e_lo_10")



df_fd$fd_ei_up_1 <- df_fd$fd_e_up_1 / df_fd$fd_i_up_1
df_fd$fd_ei_up_10 <- df_fd$fd_e_up_10 / df_fd$fd_i_up_10

df_fd$fd_ei_lo_1 <- df_fd$fd_e_lo_1 / df_fd$fd_i_lo_1
df_fd$fd_ei_lo_10 <- df_fd$fd_e_lo_10 / df_fd$fd_i_lo_10


df_fd$fd_i_up <- (df_fd$fd_i_up_1 + df_fd$fd_i_up_10)/2
df_fd$fd_i_lo <- (df_fd$fd_i_lo_1 + df_fd$fd_i_lo_10)/2

df_fd$fd_e_up <- (df_fd$fd_e_up_1 + df_fd$fd_e_up_10)/2
df_fd$fd_e_lo <- (df_fd$fd_e_lo_1 + df_fd$fd_e_lo_10)/2

df_fd$fd_ei_up <- (df_fd$fd_ei_up_1+df_fd$fd_ei_up_10)/2
df_fd$fd_ei_lo <- (df_fd$fd_ei_lo_1+df_fd$fd_ei_lo_10)/2

df_fd$fd_i_delta <- df_fd$fd_i_up - df_fd$fd_i_lo  
df_fd$fd_e_delta <- df_fd$fd_e_up - df_fd$fd_e_lo 
df_fd$fd_ei_delta <- df_fd$fd_ei_up - df_fd$fd_ei_lo 

df_fd$fd_i_av <- (df_fd$fd_i_up + df_fd$fd_i_lo) /2
df_fd$fd_e_av <- (df_fd$fd_e_up + df_fd$fd_e_lo) /2
df_fd$fd_ei_av <- (df_fd$fd_ei_up + df_fd$fd_ei_lo) /2

df_fd$fd_i_bias <- (df_fd$fd_i_delta / df_fd$fd_i_av) *100
df_fd$fd_e_bias <- (df_fd$fd_e_delta / df_fd$fd_e_av) *100
df_fd$fd_ei_bias <- (df_fd$fd_ei_delta / df_fd$fd_ei_av) *100

df_fd[,grep("_1", names(df_fd))] <- NULL
df_fd[,grep("delta", names(df_fd))] <- NULL

df_fd$fd_e_bias[is.nan(df_fd$fd_e_bias)] <- 0
df_fd$fd_ei_bias[is.nan(df_fd$fd_ei_bias)] <- 0

head(df_fd)

```




Extract the appropriate MLD and clinical variables.
```{r}
df_mld <- data.frame(raw_data2$numpat,
                     raw_data2$I_MLD_H_1,
                     raw_data2$I_MLD_B_1,
                     raw_data2$E_MLD_H_1,
                     raw_data2$E_MLD_B_1,
                     raw_data2$age,
                     raw_data2$BMI,
                     raw_data2$ACQ6,
                     raw_data2$VEMSp_1,
                     raw_data2$CVFp_1,
                     raw_data2$VEMS.CVF_1,
                     raw_data2$valeur_no_ex,
                     raw_data2$valeur_no_alveo)
    

colnames(df_mld) <- c("patient",
                      "mld_i_up",
                      "mld_i_lo",
                      "mld_e_up",
                      "mld_e_lo",
                      "age",
                      "bmi",
                      "acq",
                      "fev1",
                      "fvc",
                      "f.f",
                      "feno",
                      "fano")

df_mld$mld_ei_up <- df_mld$mld_e_up/df_mld$mld_i_up
df_mld$mld_ei_lo <- df_mld$mld_e_lo/df_mld$mld_i_lo

df_mld$mld_i_delta <- df_mld$mld_i_up - df_mld$mld_i_lo 
df_mld$mld_e_delta <- df_mld$mld_e_up - df_mld$mld_e_lo 
df_mld$mld_ei_delta <- df_mld$mld_ei_up - df_mld$mld_ei_lo 

df_mld$mld_i_av <- (df_mld$mld_i_up + df_mld$mld_i_lo)/2 
df_mld$mld_e_av <- (df_mld$mld_e_up + df_mld$mld_e_lo)/2
df_mld$mld_ei_av <- (df_mld$mld_ei_up + df_mld$mld_ei_lo)/2

df_mld$mld_i_bias <- (df_mld$mld_i_delta / df_mld$mld_i_av) *100
df_mld$mld_e_bias <- (df_mld$mld_e_delta / df_mld$mld_e_av) *100
df_mld$mld_ei_bias <- (df_mld$mld_ei_delta / df_mld$mld_ei_av) *100

df_mld[,grep("delta", names(df_mld))] <- NULL

head(df_mld)

```

Do some data wrangling to make a dataframe with only those CT vars we want, in the order we want.
```{r}
df <- merge(df_mld, df_fd, by = "patient", all.x = TRUE)

df$acq <- as.numeric(as.character(df$acq))

ct <- df[, c("fd_i_up",
             "fd_i_lo",
             "fd_i_av",
             "mld_i_up",
             "mld_i_lo",
             "mld_i_av",
             "fd_e_up",
             "fd_e_lo",
             "fd_e_av",
             "mld_e_up",
             "mld_e_lo",
             "mld_e_av",
             "fd_ei_up",
             "fd_ei_lo",
             "fd_ei_av",
             "mld_ei_up",
             "mld_ei_lo",
             "mld_ei_av",
             "fd_i_bias",
             "fd_e_bias",
             "fd_ei_bias",
             "mld_i_bias",
             "mld_e_bias",
             "mld_ei_bias")]


```

Make functions for correlations and p-values.

```{r}

set.seed(123)

age_cor <- function(x) {
    age <- df$age
    k <- data.frame(age, x)
    ks <- subset(k, subset = (complete.cases(k)))
    round(cor(ks$age, ks$x, method = c("spearman")), digits = 2)
    }

age_cor_test <- function(x) {   
    age <- df$age
    k <- data.frame(age, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$age, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)
    }

bmi_cor <- function(x) {
    bmi <- df$bmi
    k <- data.frame(bmi, x)
    ks <- subset(k, subset = (complete.cases(k)))
    round(cor(ks$bmi, ks$x, method = c("spearman")), digits = 2)
}

bmi_cor_test <- function(x) {   
    bmi <- df$bmi
    k <- data.frame(bmi, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$bmi, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)
    }






fev1_cor <- function(x) {
    fev1 <- df$fev1
    k <- data.frame(fev1, x)
    ks <- subset(k, subset = (complete.cases((k))))
    round(cor(ks$fev1, ks$x, method = c("spearman")), digits=2)}

fev1_cor_test <- function(x) {   
    fev1 <- df$fev1
    k <- data.frame(fev1, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$fev1, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)}



fvc_cor <- function(x) {
    fvc <- df$fvc
    k <- data.frame(fvc, x)
    ks <- subset(k, subset = (complete.cases((k))))
    round(cor(ks$fvc, ks$x, method = c("spearman")), digits=2)}

fvc_cor_test <- function(x) {   
    fvc <- df$fvc
    k <- data.frame(fvc, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$fvc, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)}

f.f_cor <- function(x) {
    f.f <- df$f.f
    k <- data.frame(f.f, x)
    ks <- subset(k, subset = (complete.cases((k))))
    round(cor(ks$f.f, ks$x, method = c("spearman")), digits=2)}

f.f_cor_test <- function(x) {   
    f.f <- df$f.f
    k <- data.frame(f.f, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$f.f, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)}

feno_cor <- function(x) {
    feno <- df$feno
    k <- data.frame(feno, x)
    ks <- subset(k, subset = (complete.cases((k))))
    round(cor(ks$feno, ks$x, method = c("spearman")), digits=2)}

feno_cor_test <- function(x) {   
    feno <- df$feno
    k <- data.frame(feno, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$feno, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)}



fano_cor <- function(x) {
    fano <- df$fano
    k <- data.frame(fano, x)
    ks <- subset(k, subset = (complete.cases((k))))
    round(cor(ks$fano, ks$x, method = c("spearman")), digits=2)}

fano_cor_test <- function(x) {   
    fano <- df$fano
    k <- data.frame(fano, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$fano, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)}


acq_cor <- function(x) {
    acq <- df$acq
    k <- data.frame(acq, x)
    ks <- subset(k, subset = (complete.cases((k))))
    round(cor(ks$acq, ks$x, method = c("spearman")), digits=2)}

acq_cor_test <- function(x) {   
    acq <- df$acq
    k <- data.frame(acq, x)
    ks <- subset(k, subset = (complete.cases((k))))
    ct <- cor.test(ks$acq, ks$x, method = c("spearman"))
    round(ct[[3]], digits=3)}
```

Apply it to all ct variables.
```{r}

corr_coeff_age <- sapply(ct, age_cor)
corr_p_with_age <- sapply(ct, age_cor_test)

corr_coeff_bmi <- sapply(ct, bmi_cor)
corr_p_with_bmi <- sapply(ct, bmi_cor_test)

corr_coeff_fev1 <- sapply(ct, fev1_cor)
corr_p_with_fev1 <- sapply(ct, fev1_cor_test)

corr_coeff_fvc <- sapply(ct, fvc_cor)
corr_p_with_fvc <- sapply(ct, fvc_cor_test)

corr_coeff_f.f <- sapply(ct, f.f_cor)
corr_p_with_f.f <- sapply(ct, f.f_cor_test)

corr_coeff_feno <- sapply(ct, feno_cor)
corr_p_with_feno <- sapply(ct, feno_cor_test)

corr_coeff_fano <- sapply(ct, fano_cor)
corr_p_with_fano <- sapply(ct, fano_cor_test)

corr_coeff_acq <- sapply(ct, acq_cor)
corr_p_with_acq <- sapply(ct, acq_cor_test)
```

Make a summary table.
```{r}
correlations <- data.frame(corr_coeff_age, 
                           corr_p_with_age,
                           corr_coeff_bmi, 
                           corr_p_with_bmi,
                           corr_coeff_acq, 
                           corr_p_with_acq,
                           corr_coeff_fev1, 
                           corr_p_with_fev1,
                           corr_coeff_fvc,
                           corr_p_with_fvc,
                           corr_coeff_f.f, 
                           corr_p_with_f.f,
                           corr_coeff_feno, 
                           corr_p_with_feno,
                           corr_coeff_fano, 
                           corr_p_with_fano)

write.csv(correlations, file = "C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/Submission2_Investigative_Radiology/stats-tables/clinical_correlations.csv")
write.csv2(correlations, file = "C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/Submission2_Investigative_Radiology/stats-tables/clinical_correlations_csv2.csv")
```





